Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-19 paper/main.tex
3 days
reply to comments

%% Revision documents
\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
% FIXME: \sout is not very good. It cannot handle commands within its argument.
\newcommand{\SuggestEdit}[3][red]{\textcolor{#1}{\sout{#2}}\textcolor{#1}{#3}}
\let\svthefootnote\thefootnote
\newcommand\colorfootnote[2][black]{\def\thefootnote{\color{#1}\svthefootnote}%
  \footnote{\color{#1}#2}\def\thefootnote{\color{black}\svthefootnote}}
\newcommand\RevComment[3][red]{\protect\colorfootnote[#1]{{\textbf{[#2: #3]}}}}
% Command for edits, command for commenting and color
\newcommand\newrevisor[3]{%%
  \colorlet{#1}{#3}
  \expandafter\newcommand\csname#1\endcsname[2]{\SuggestEdit[#1]{##1}{##2}}%%
  \expandafter\newcommand\csname#2\endcsname[1]{\RevComment[#1]{#2}{##1}}%%
}
\newrevisor{manuel}{MANUEL}{Purple}

@@

Ekhine: How many runs do we do? 10 20? 10, cambio el texto

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs on the same problem instances.  As an example, Fig.~\ref{fig:kra32} \textcolor{red}{por qué no mepezamos con la fig1?}
shows the mean fitness of the solutions evaluated by uMM and CEGO on one QAP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, the plot shows that \textcolor{red}{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}
the fitness of the solutions does not improve monotonically and new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

When we compare the behaviour of CEGO and uMM on synthetic LOP instances
(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
uniformity of the instances. In particular, CEGO starts with a random sampling
of X solutions, with are quite poor as expected, followed by building and
updating the Kriging model based new evaluations, which leads to an impressive
improvement in less than 50 evaluations. However, progress is more or less
halted at this point and the model seems to have trouble generating better
solutions. Lower uniformity values (e.g., $\mu=5$) show worsening evaluations
despite the increase in solutions used to build the model.

On the other hand, uMM converges much more slowly than CEGO, only reaching the
same fitness values around evaluation 250. Yet, uMM is able to keep the rate of
improvement and completely overtake CEGO around evaluation 350. As in the case
of CEGO, lower uniformity values lead to worse performance, although the
difference is much smaller for uMM than it was for CEGO.

In summary, the Kriging model clearly helps CEGO to quickly identify good
solutions in very few evaluations. However, for some reason, CEGO is not able to
further improve those solutions given more data. This may due to the
model not being able to produce accurate predictions or the underlying
optimizer not being able to find improved solutions using the model.\MANUEL{perhaps we should measure the prediction error and see if it grows or decreases. That would answer this question.}

@@

%: LOP reales: he puesto 3 de diferentes tamanos. Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
  \end{figure}

The fast initial convergence of CEGO for synthetic LOP instances does not seem
to translate to the instances available in LOPLIB~\citep{}. We show in
Fig.~\ref{fig:loplib} three examples with different permutation size, however,
the results are consistent for other LOPLIB instances (complete results
available as supplementary material \supplement). Due to extremely long
runtimes of CEGO and constraints in our computing system, we set a maximum
wall-clock time of 5 days for each run of CEGO. On these instances, the
behaviour of CEGO and uMM is much more similar (at least up to the termination
point of CEGO). It appears that our synthetic LOP instances have a fitness
landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
instances do not present such landscape. Comparing CEGO and uMM results, the
differences get smaller, in favour of uMM, with larger instance size. Our
conjecture is that building an accurate Bayesian model and searching for good
solutions on it becomes harder for larger fitness landscapes. 
\textcolor{red}{Parece que el paper va del CEGO en vez del UMM :p. Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}

@@

\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
\end{figure}

The situation is very similar for the QAP, where both algorithms show the same
behaviour as in the PFSP for all instances tested, except for
\texttt{kra32}. As shown in Fig~\ref{fig:kra32}, in this instance both
algorithms are able to improve over the number of evaluations, although not
with ease. As in the synthetic LOP instances, we again observe that CEGO does
improve faster than UMM initially, yet, around evaluation 250, it appears to
get stuck and around evaluation 300, UMM overtakes it. We believe the reason is
the same as before, that is, the model is not able to adequately reflect the
ruggedness of the actual landscape and, hence, the best solutions found by the
GA searching the model are not improved solutions for the actual problem. Interestingly, \texttt{kra32} is rectangular
whereas \texttt{tho32} contains rectangular distances and \texttt{nug12} and \texttt{nug30} contain Manhattan distances of rectangular grids.\MANUEL{so they all have the same distance type!!!}

>>>>

%% Revision documents
\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
% FIXME: \sout is not very good. It cannot handle commands within its argument.
\newcommand{\SuggestEdit}[3][red]{\textcolor{#1}{\sout{#2}}\textcolor{#1}{#3}}
\let\svthefootnote\thefootnote
\newcommand\colorfootnote[2][black]{\def\thefootnote{\color{#1}\svthefootnote}%
  \footnote{\color{#1}#2}\def\thefootnote{\color{black}\svthefootnote}}
\newcommand\RevComment[3][red]{\protect\colorfootnote[#1]{{\textbf{[#2: #3]}}}}
% Command for edits, command for commenting and color
\newcommand\newrevisor[3]{%%
  \colorlet{#1}{#3}
  \expandafter\newcommand\csname#1\endcsname[2]{\SuggestEdit[#1]{##1}{##2}}%%
  \expandafter\newcommand\csname#2\endcsname[1]{\RevComment[#1]{#2}{##1}}%%
}
\newrevisor{manuel}{MANUEL}{Purple}
\newrevisor{ekhine}{EKHINE}{red}

@@

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs on the same problem instances.  As an example, each line in Fig.~\ref{fig:lop_synth} 
shows the mean fitness of the solutions evaluated by either uMM or CEGO on one synthetic LOP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, it is often the case that  \EKHINE{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}\MANUEL{Así está mejor? Quiero llamar la atención porque normalmente se muestra la evolución de la mejor solución encontrada pero nosotros estamos mostrando el fitness de cada solución evaluada}
% the fitness of the solutions does not improve monotonically and
new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

When we compare the behaviour of CEGO and uMM on synthetic LOP instances
(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
uniformity of the instances. In particular, CEGO starts with a random sampling
of X solutions, with are quite poor as expected, followed by building and
updating the Kriging model based new evaluations, which leads to an impressive
improvement in less than 50 evaluations. However, progress is more or less
halted at this point and the model seems to have trouble generating better
solutions. Lower uniformity values (e.g., $\mu=5$) show worsening evaluations
despite the increase in solutions used to build the model. In summary, the
Kriging model clearly helps CEGO to quickly identify good solutions in very few
evaluations. However, for some reason, CEGO is not able to further improve
those solutions given more data. This may due to the model not being able to
produce accurate predictions or the underlying optimizer not being able to find
improved solutions using the model.\MANUEL{perhaps we should measure the
  prediction error and see if it grows or decreases. That would answer this
  question.}

On the other hand, uMM converges much more slowly than CEGO, only reaching the
same fitness values around evaluation 250. Yet, uMM is able to keep the rate of
improvement and completely overtake CEGO around evaluation 350. As in the case
of CEGO, lower uniformity values lead to worse performance, although the
difference is much smaller for uMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}

@@

%: LOP reales: he puesto 3 de diferentes tamanos. Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
\begin{figure}[tb]
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
  \end{figure}

Results are quite different on the 
%The fast initial convergence of CEGO for synthetic LOP instances does not seem
%to translate to the
instances available in LOPLIB~\citep{}. We show in
Fig.~\ref{fig:loplib} three examples with different permutation size, however,
the results are consistent for other LOPLIB instances (complete results
available as supplementary material \supplement). Due to extremely long
runtimes of CEGO and constraints in our computing system, we set a maximum
wall-clock time of 3 days for each run of CEGO. On these instances, the
behaviour of CEGO and uMM is much more similar (at least up to the termination
point of CEGO). It appears that our synthetic LOP instances have a fitness
landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
instances do not present such landscape. Comparing CEGO and uMM results, the
differences get smaller, in favour of uMM, with larger instance size. Our
conjecture is that building an accurate Bayesian model and searching for good
solutions on it becomes harder for larger fitness landscapes.
\EKHINE{Parece que el paper va del CEGO en vez del UMM :p. Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}

@@

\begin{figure}[tb]
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
\end{figure}

The situation is very similar for the QAP, where both algorithms show the same
behaviour as in the PFSP for all instances tested, except for
\texttt{kra32}. As shown in Fig~\ref{fig:kra32}, in this instance both
algorithms are able to improve over the number of evaluations, although not
with ease. As in the synthetic LOP instances, we again observe that CEGO does
improve faster than UMM initially, yet, around evaluation 250, it appears to
get stuck and around evaluation 300, UMM overtakes it. We believe the reason is
the same as before, that is, the model is not able to adequately reflect the
ruggedness of the actual landscape and, hence, the best solutions found by the
GA searching the model are not improved solutions for the actual problem. Interestingly, \texttt{kra32} is rectangular
whereas \texttt{tho32} contains rectangular distances and \texttt{nug12} and \texttt{nug30} contain Manhattan distances of rectangular grids.\MANUEL{so they all have the same distance type!!! What is the difference then? Ask Thomas?}
